### 2025-10-31 write-ups/manuscript/kid-tangrams-ms.Rmd
figs, intro, cites

---
title             : |
  | Preschoolers can form conventional pacts with each other 
  | to communicate about novel referents

author: 
  - name          : "Veronica Boyce"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "vboyce@stanford.edu"
  - name          : "Robert Z. Sparks"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

@@

abstract: > 
  Learning language requires learning not only the content of language, but also how to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents. Some early experiments are interpreted as evidence that 4-5-year-old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game. Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm. Across TODO pairs of children, we found that 4-5-year-olds successfully established reference with each other. Children were TODO% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.
  <!-- https://tinyurl.com/ybremelq -->

@@

TODO all figures could be smaller!

@@

Iterated reference games provide an important paradigm for studying referential communication. In these games, one player repeatedly describes a set of abstract shapes to a partner so they can identify the target images [@clark1986; @krauss1964; @hawkins2020b; @boyce2024]. Over repetition, features of the initial descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image. Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform later referring expressions. 
Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.

@@

TODO update diagram with expt 3 stim.

```{r interface, fig.env = "figure", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure. Panel A shows the experimental setup. Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only. Panel C shows the procedure for Experiment 1; within critical blocks targets were ordered randomly. Panel D shows the procedure for Experiment 2. \\label{game}", cache=FALSE}

@@

# edit diagram at 
# https://docs.google.com/drawings/d/1gio1CNSedECJGrkrNDUi_AydPZcYzESfl4WhZD2l9Bc/edit?usp=sharing
```

@@

More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively. 
* [@sanjuan2025] When 4-5 year old children learn that some objects are only visible to them and not their partner, they are faster to rule out these competitors and select a target than when they believe the interlocuter is aware of the competitors as well. 
* [@nilsen2009] take into account partners perspective for both speaker and listening (4-5 yos); 3-4 year olds also on listening only; relate this to inhibitory control.

@@

A common thread among many developmental studies is that children perform better when the cognitive load of the task is reduced, i.e. when there are fewer possible referents [@abbot-smith2016; @sanjuan2015].  These findings paired with evidence of emerging communicative skills in young children suggest that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.

Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game. In one study, 8-10-year-olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance and highly variable between dyads [@branigan2016]. 4-6-year-olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019]. 4-8-year-old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024]. In this task, even 4-year-olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024]. Together, these findings provide evidence of young children’s ability to communicate about novel referents under the right conditions.

- Bohn, Kachel, Tomasello (2019) Young Children spontaneously recreate core properties of language in a new modality

    - 4, 6, (and 8 in a separate thing) communicating target image via video link (no audio after practice trials)

    - comprehension good and increasing, gestures used back by second partner when roles switch

    - (out of 5 pictures, including “blank”)

    - do see more similaritiy w/i dyads especially for 6 versus 4 yos

May want to set up for change over age? And a little more on why 4-5 is a good age -- it's a time of rapid development, we think from prior work that children can comprehend the task. Just before big changes in linguistic input from school + reading. ...

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across three studies including a total of ZZ pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

### Materials
For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B). We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on tablets that were locked in a kiosk mode so children could not navigate away from the game.

@@

```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time. Error bars are bootstrapped 95% CIs with a linear trend line overlaid. \\label{acc}" }
joint_data |>
  ggplot(aes(x = trial+1, y = correct)) +
    geom_smooth(method = "lm", color = "black") +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot") +
  #stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  labs(color="Block")+
  xlab("Trial") +
  ylab("Accuracy") +
  geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_x_continuous(breaks = seq(1, 16, 1), expand = c(0.02, 0.02)) +
  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
  coord_cartesian(ylim = c(0.3,1))+
  geom_hline(yintercept = .5) +
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of description produced by the teller each trial. Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs. \\label{length}" }
joint_data |> 
  ggplot(aes(x = trial+1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
    geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_color_manual(values = time_palette) +
  geom_smooth(method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020b; @boyce2024]. We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced.^[`r description_form |> form()`] On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words, and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}). Thus, children's increasing speed was not from shorter utterances, but instead from some combination of improved task understanding, faster utterance planning, and faster decisions of what to select.

@@

```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources. Dots are means and lines are bootstrapped 95% CIs.  \\label{sim}" }

@@

TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3

```{=latex}
\begin{table}
	\caption{Example descriptions children successfully used to identify different target images in Experiment 2.}
	\label{example}
	\begin{tabular}[t]{p{16em}p{4em}}
		\hline
		%hold
		$\bullet$ person & 		\multirow{6}{*}{\includegraphics[width=.5in]{I1.jpg}}\\

@@

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
joint_similarities |> 
  filter(game1==game2) |> 
  filter(later==3) |> 
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>left_join(labels) |> 
  ggplot(aes(earlier+1, sim, color = img_source)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line") +
    geom_smooth(color="black", method="lm")+
  facet_grid(str_c("Expt ", expt)~same_speaker) +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
scale_color_brewer(palette="Dark2") +
  theme(
    legend.position = "right",
    legend.text = element_markdown(color = "black", size = 11),
    panel.grid = element_blank(),
    strip.background = element_blank()
  )

@@

We did not account for children's age and gender when pairing children to play the game, so children of different ages (within the 4-5 year old target age range) sometimes played together, and children sometimes played with same-gender and sometimes with different-gender partners. As children of this age are rapidly developing their linguistic skills (CITATION NEEDED), we looked at possible effects of children's age and gender on performance as an exploratory measure.

```{r age-acc, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Accuracy of selections across all experiments, graphed by the mean age of child in that game, and colored by the number of girls in the pair. Each dot is a game, and linear fits are overlayed. \\label{ageacc}" }
joint_data |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
  ggplot(aes(x = mean_age, y = correct, color = as.character(girlness))) +
  labs(x="Mean age of participants (months)", y="Accuracy", color="Number of girls")+
  geom_jitter(data=joint_data|>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(correct=mean(correct, na.rm=T))) +
  scale_color_manual(values=c("#DC322FFF","#2AA198FF" ,"#268BD2FF"  ))+
  geom_smooth(method = "lm", show.legend=F)
```

```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender. Each dot is a game, and linear fits are overlayed. \\label{agewords}" }
joint_data |>  ggplot(aes(x = speaker_age, y = words, color = speaker_gender)) +
  geom_smooth(method = "lm", show.legend=F)+
  geom_jitter(data=joint_data |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))+
  scale_color_manual(values=c("#268BD2FF", "#DC322FFF"))+
  labs(x="Teller Age (months)", y="Words", color="Teller gender")
```

@@

It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults. In this case, the low number of options and relatively easy-to-describe shapes may have obviated the need for long initial descriptions. Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024]. However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification. Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.

>>>>

---
title             : |
  | Preschoolers can form conventional pacts with each other 
  | to communicate about novel referents (TODO?)
shorttitle: Preschoolers can form conventional pacts (TODO?)
author: 
  - name          : "Veronica Boyce"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "vboyce@stanford.edu"
  - name          : "Robert Z. Sparks"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

@@

abstract: > 
  Learning language requires learning not only the content of language, but also how to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents. Some early experiments are interpreted as evidence that 4-5-year-old children cannot generate the mutually understandable names needed to succeed in an iterated reference game. Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm. Across TODO pairs of children, we found that 4-5-year-olds successfully established reference with each other. Children were TODO% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.
  <!-- https://tinyurl.com/ybremelq -->

@@

@@

Iterated reference games provide an important paradigm for studying referential communication. In these games, one player repeatedly describes a set of abstract shapes to a partner so they can identify the target images [@clark1986; @krauss1964; @hawkins2020dynamics; @boyce2024]. Over repetition, features of the initial descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image. Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform later referring expressions. 
Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.

@@

```{r interface, fig.env = "figure", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure. Panel A shows the experimental setup. Panel B shows the procedures for each experiment;  within critical blocks targets were ordered randomly.\\label{game}", cache=FALSE}

@@

# edit diagram at 
# https://docs.google.com/drawings/d/1kLfIGdPGeEAQgXG--FE7eyLKzX0DjTZxwfQOIhvW19I/edit?usp=sharing
```

@@

More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2013; @matthews2010]. When a change in visual context renders previous descriptions overinformative, 6-year-olds (but not 4-year-olds) are more likely to use consistent referential expressions when they keep talking to the same partner, but switch to simpler terms with a new partner [@koymen2014] These studies suggest that young children are sensitive to the norm of consistent descriptions in cooperative communication.

Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. 4-5 year old children are faster to rule out competitors and select the correct target when they have learned that their partner is not aware of the competitor object. Children's ability to take partner's perspectives and knowledge states into account is evolving over the preschool years, but overall, by late preschool, children can reason about others’ perspectives to communicate more effectively.

@@

A common thread among many developmental studies is that children perform better when the cognitive load of the task is reduced, i.e. when there are fewer possible referents [@abbot-smith2015; @sanjuan2015].  These findings paired with evidence of emerging communicative skills in young children suggest that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.

Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game. In one study, 8-10-year-olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance and highly variable between dyads [@branigan2016]. In an iterated reference game where 4- and 6- year-olds communicated to each other via gestures, children were successful at communicating a repertoire of 5 image options with each other [@bohn2019]. Children's comprehension increased over repetitions, and when children switched roles, they often used their partner's gestures back at each other. The 6-year-olds showed greater within dyad similarity in gestures than the 4-year-olds, indicating that children's abilities to form conventional pacts is increasing in this age range.

In an iterated reference game using natural language to communicate, 4-8-year-old children succeeded when playing with a parent using a simple, child-friendly tablet-based task [@leung2024]. In this task, even 4-year-olds had an initial accuracy above 80% on two-way choice, which rose to above 90% in later repetitions [@leung2024]. Together, the recent work on both iterated and non-iterated reference games with young children provides evidence of young children’s ability to communicate with each other and with adults under the right conditions. The late preschool years are a time of rapid development for children as a number of communicative skills are improving between the ages of 3 and 6 [TODO maybe more citations, or re-citations].

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across three studies including a total of ZZ pairs of 4-5-year-old preschool children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

### Materials
For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B). We coded the matching game using Empirica [@almaatouq2021], hosted it on a server, and then accessed the game on tablets that were locked in a kiosk mode so children could not navigate away from the game.

@@

```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time. Error bars are bootstrapped 95% CIs with a linear trend line overlaid. \\label{acc}" }
joint_data |>
  ggplot(aes(x = trial+1, y = correct)) +
    geom_smooth(method = "lm", color = "black") +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot") +
  #stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  labs(color="Block")+
  xlab("Trial") +
  ylab("Accuracy") +
  geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_x_continuous(breaks = seq(1, 16, 1), expand = c(0.02, 0.02)) +
  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
  coord_cartesian(ylim = c(0.3,1))+
  geom_hline(yintercept = .5) +
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Length of description produced by the teller each trial. Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs. \\label{length}" }
joint_data |> 
  ggplot(aes(x = trial+1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
    geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_color_manual(values = time_palette) +
  geom_smooth(method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

In iterated reference games with adults, description lengths usually shorten over repeated references [@clark1986; @hawkins2020dynamics; @boyce2024]. We were curious if children's descriptions would display the same trend, so we ran a Bayesian mixed effects linear regression predicting the number of words in the description the "teller" produced.^[`r description_form |> form()`] On the first critical trial, descriptions averaged `r stats_text(description_mod_1, 1)` words, and description length was relatively stable over time (change of `r stats_text(description_mod_1, 2)` words per trial, Figure \ref{length}). Thus, children's increasing speed was not from shorter utterances, but instead from some combination of improved task understanding, faster utterance planning, and faster decisions of what to select.

@@

```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources. Dots are means and lines are bootstrapped 95% CIs.  \\label{sim}" }

@@

```{=latex}
\begin{table}
	\caption{Example descriptions children successfully used to identify different target images in Experiment 2. TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3
}
	\label{example}
	\begin{tabular}[t]{p{16em}p{4em}}
		\hline
		%hold
		$\bullet$ person & 		\multirow{6}{*}{\includegraphics[width=.5in]{I1.jpg}}\\

@@

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
joint_similarities |> 
  filter(game1==game2) |> 
  filter(later==3) |> 
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>left_join(labels) |> 
  ggplot(aes(earlier+1, sim, color = img_source)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line") +
    geom_smooth(color="black", method="lm")+
  facet_grid(str_c("Expt ", expt)~same_speaker) +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
scale_color_brewer(palette="Dark2") +
  theme(
    legend.position = "right",
    legend.text = element_markdown(color = "black", size = 11),
    panel.grid = element_blank(),
    strip.background = element_blank()
  )

@@

We did not account for children's age and gender when pairing children to play the game, so children of different ages (within the 4-5 year old target age range) sometimes played together, and children sometimes played with same-gender and sometimes with different-gender partners. As children of this age are rapidly developing their linguistic skills (TODO CITATION NEEDED), we looked at possible effects of children's age and gender on performance as an exploratory measure.

```{r age-acc, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=2, fig.cap = "Accuracy of selections across all experiments, graphed by the mean age of child in that game, and colored by the number of girls in the pair. Each dot is a game, and linear fits are overlayed. \\label{ageacc}" }
joint_data |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
  ggplot(aes(x = mean_age, y = correct, color = as.character(girlness))) +
  labs(x="Mean age of participants (months)", y="Accuracy", color="Number of girls")+
  geom_jitter(data=joint_data|>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(correct=mean(correct, na.rm=T))) +
  scale_x_continuous(breaks = seq(1, 16, 1), expand = c(0.02, 0.02)) +
  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
  coord_cartesian(ylim = c(0.3,1))+
  geom_hline(yintercept = .5) +
  scale_color_manual(values=c("#DC322FFF","#2AA198FF" ,"#268BD2FF"  ))+
  geom_smooth(method = "lm", show.legend=F)
```

```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=2, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender. Each dot is a game, and linear fits are overlayed. \\label{agewords}" }
joint_data |>  ggplot(aes(x = speaker_age, y = words, color = speaker_gender)) +
  geom_smooth(method = "lm", show.legend=F)+
  geom_jitter(data=joint_data |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))+
  scale_color_manual(values=c("#268BD2FF", "#DC322FFF"))+
  labs(x="Teller Age (months)", y="Words", color="Teller gender")
```

@@

It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults. In this case, the low number of options and relatively easy-to-describe shapes may have obviated the need for long initial descriptions. Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020dynamics; @boyce2024]. However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification. Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.
